---
title: "Class 11: Structural Bioinformatics (Part I)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Revisit PDB

Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and determine the percentage of structures solved by X-Ray and Electron Microscopy. Also can you determine what proportion of structures are protein? 

Download a CSV file from: http://www.rcsb.org/stats/summary on 2019-05-08

```{r}
db <- read.csv("Data Export Summary.csv", row.names= 1)
head(db)
```

```{r}
## Total number of structures in the PDB database
sum(db$Total)

## Percentages of structures solved by X-Ray and Electron Microscopy
(db$Total)/(sum(db$Total))*100
```
89.3 % are solved by X-ray; 2.1% are solved by Electron Microscopy. 



What percent are protein?
```{r}
(sum(db$Proteins) / sum(db$Total))*100
```
92. 76%



Q2: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

1157 proteins



## Section 3. Using Bio3D

Let's read the PDB file we downloaded into R to see how things ive in this place 

```{r}
library(bio3d)
```


```{r}
pdb <- read.pdb("1hsg.pdb")
pdb
```


```{r}
aa321(pdb$seqres)
```



```{r}
pdb$atom
```



##Atom selection examples

Let's select residue 1
```{r}
inds <- atom.select(pdb, resno = 10)
```




```{r}
pdb$atom[inds$atom,]
```


### Make a protein only and ligand only PDB file

Here we need to select our protein with `atom.select()` then use `write.pdb()`

```{r}
prot <- atom.select(pdb, "protein", value = TRUE)
prot
```

```{r}
write.pdb(prot, file="1hsg_protein")
```

```{r}
lig <- atom.select(pdb, "ligand", value=TRUE)
lig
```

```{r}
write.pdb(lig, file="1hsg_ligand.pdb")
```

```{r}
#install.packages("rgl")
#install.packages("xquart2")

```

```{r}
#install.packages("devtools")
```


```{r}
devtools::install_bitbucket("Grantlab/bio3d-view")
library("bio3d.view")

```
```{r}
view(pdb,"overview", col="sse")
```

```{r}
view(pdb, "overview", col="sse")
```


```{r}
pdbex <- read.pdb("1he1")
```

```{r}
modes <- nma(pdbex)
m7 <- mktrj(modes, mode=7, file="mode_7.pdb")
view(m7, col=vec2color(rmsf(m7)))
```


Download some example PDB files
```{r}
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split=TRUE)
```


```{r}
pdbs <- pdbaln(files, fit = TRUE)
pdbs
```



```{r}
pc.xray <- pca(pdbs)
```


```{r}
plot(pc.xray)
```



```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```





