---
title: "Structural Bioinformatics II"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Obtain and prepare our structure for "docking"

Here we download the HIV-Ps structure from the PDB database and extract out the protein and ligand portions to separate new PDB format files. 
```{r}
library(bio3d)
file <- get.pdb("1hsg")
hiv <- read.pdb(file)
hiv


```

```{r}
# Select protein only
prot <- atom.select(hiv, "protein", value=TRUE)
prot
write.pdb(prot, file = "1hsg_protein.pdb" )

```


```{r}
# Select ligand only
lig <- atom.select(hiv, "ligand", value = TRUE)
lig
write.pdb(lig, file="1hsg_ligand.pdb")
```



## Add Hydrogens and charges in ADT

We opened our protein only PDB file in AutoDocTools and added hydrogens and atom-types needed for docking calculations

## Run docking

We will use AutoDoc Vina here at the UNIX command line! 


```{r}
res <- read.pdb("all.pdbqt", multi = TRUE)
res
```

```{r}
write.pdb(res, "results.pdb")
```


Compare our docking poses to the MERK drug structure

```{r}
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)
```

- the smaller the value here, the smaller the distance between the tested ligand and the MERK drug 



## Normal Mode Analysis example

```{r}
pdb <- read.pdb("1hel")
modes <- nma(pdb)
m7 <- mktrj(modes, mode=7, file="mode_7.pdb")
```


```{r}
plot(modes, sse=pdb)
```























































