---
title: "Section_1B_worksheet"
author: "Brie Diaz"
date: "4/19/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("bio3d")
```

## Improve the analysis code 

```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

##Q1 What type of object is returned from the read.pdb() function?

```{r}
is.list(s1)
```
- the read.pdb() function reads a PDB file and returns a list 


##Q2 What does the trim.pbd() function do?

```{r}
?trim.pdb
```
- Trims a PBD object to a subset of atoms; description: produces a new smaller PDB object, containing a subset of atoms, from a given larger PDB object 
 
 
##Q3 What input parameter would turn off the marginal black and gray rectangles in the plots and what do they represent in this case? 

- remove sse or set to FALSE; also, you could set both top = FALSE and bot = FALSE 

```{r}
?plot.bio3d
```

plotb3 draws a standard scatter plot with optional secondary structure in the marginal regions







##Q5 Which proteins are more similar to each other in their B-factor trends
```{r}
hc <- hclust(dist(rbind(s1.b, s2.b, s3.b)))
plot(hc)
```

 - Proteins 1AKE and 1E4Y (represented by s2.b and s3.b, respectively) are more similar to each other


## Q6 How would you generalize the original code above to work with any set of input protein structures?


### Don't ever start like this: copy and paste each line of function
makeplot <- function(x) {
s1 <- read.pdb(x) # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
}


### instead, make small working snippets of code
First, access the proteins sequence from PDP and save as "pseq"
```{r}
x <- "4AKE"
pseq <- read.pdb(x)
pseq

```


Next, trim protein sequence so that it includes only chain A of the protein. 
```{r}
pseq.chainA <- trim.pdb(pseq, chain="A", elety="CA")
pseq.chainA
```

Next, access the Temperature factor, represented by the "b" vector within the "atom" data frame for this trimmed sequence 

```{r}
pseq.chainA <- trim.pdb(pseq, chain="A", elety="CA")
pseq.b <- pseq.chainA$atom$b

```


Make a line plot of the Temperature Factor data with the y-axis labeled "Bfactor"


```{r}
pseq.chainA <- trim.pdb(pseq, chain="A", elety="CA")
pseq.b <- pseq.chainA$atom$b
plotb3(pseq.b, sse=pseq.chainA, type = "l", ylab="Bfactor")
```



Finally, turn overall working snippets into a function!

- The input(x) is a name of a protein sequence in PDB; simply set x = "*name_of_PDB_seq*" in the function itself

- the function will:
1. read the PDF sequence you assign to it
2. trim your protein sequence to include chain A alpha carbons; creates a new PDB object based on a selection of atoms
3. access the B factor (Temperature Factor) for all atoms in the trimmed protein sequence
4. produce a line plot of Residue vs. B Factor --> this is the overall output of the function!



```{r}
makeplot <- function(x) {
  pseq <- read.pdb(x)
  pseq.chainA <- trim.pdb(pseq, chain="A", elety="CA")
  pseq.b <- pseq.chainA$atom$b
  plotb3(pseq.b, sse=pseq.chainA, type = "l", ylab="Bfactor")
}
```


Test out new function: 

```{r}
makeplot("4AKE")
makeplot("1AKE")
makeplot("1E4Y")
```













