---
title: "Class 7: R functions and packages"
author: "Brie Diaz"
date: "4/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## More on function writing

First we will revisit our function from last day

```{r}
source("http://tinyurl.com/rescale-R")
```


Rest the **rescale()** function

```{r}
rescale(1:10)
```



##rescale(c(1:10, "string"))
 -> error


#cmd + option + i to open up an R coding bar 
```{r}
x <- c(1:10, "string")
!is.numeric(x)
```

```{r}
##rescale2(x)
```



## Function practice

Write a function to identify NA elements in two vectors

Start with a simple example input where I now what the answer should be. 
```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```


```{r}
is.na(x)
is.na(y)
```

```{r}
which(is.na(x))
which(is.na(y))
```

I am looking for the positions where it is TRUE in both vectors....
```{r}
is.na(x) & is.na(y)
```

Take the sum to find how many TRUEs there are (TRUE = 1; FALSE = 0)

```{r}
sum(is.na(x) & is.na(y))
```

This is my working snippet of code that I can use as the body of my first function!!

```{r}
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}
```


```{r}
both_na(x,y)
```

```{r}
both_na(c(NA, NA, NA), c(NA, NA, 1))
```


```{r}
both_na(c(NA, NA, NA), c(1, NA, NA, NA))
```


```{r}
both_na(c(NA, NA, NA), c(1, NA, NA, NA, NA, NA))
```



Check that the length of our inputs are equal 

```{r}
x <- c(NA, NA, NA)
y <- c(1, NA, NA, NA, NA, NA)
length(x) != length(y)

```

```{r}
both_na2 <- function(x, y) {
 if(length(x) != length(y)) {
 stop("Input x and y should be the same length")
 }
 sum( is.na(x) & is.na(y) )
}
```



Try the both_na3() function with extra features
```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
both_na3(x, y)
```


The which() function finds which positions are TRUE 
```{r}
which(c(F, F, T, F, T))
```

Write a function grade() to determine an overall grade from a vector of student homework assignment scores, dropping the lowest singe assigment score

```{r}
s1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
sum(s1)
sum(s1) - min(s1)
length(s1) -1
(sum(s1) - min(s1))/(length(s1) -1)
```


Find the minimum score
```{r}
min(s1)
```

Subtract the minimum score from the sum of the vector
```{r}
sum(s1) - min(s1)
```

Divide by the length of the vector -1 (n-1)

```{r}
(sum(s1) - min(s1))/(length(s1)-1)
```


I got my snippet, now to turn it into my first version of the grade() function
```{r}
grade <- function(x) {
  (sum(x, na.rm =TRUE) - min(x, na.rm =TRUE))/(length(x)-1)
}
```


OR, if we want the user to be able to determine what to do with NA:
- because we need to pass the na.rm through the function/s that is/are actually doing the work! 
```{r}
grade1 <- function(x, na.rm = TRUE) {
  if (na.rm) {
    return((sum(x, na.rm =TRUE) - min(x, na.rm =TRUE))/(length(x)-1))
  } else {
    (sum(x) - min(x))/(length(x)-1)
  }
}
```


```{r}
s1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
s2 <- c(100, NA, 90, 90, 90, 90, 97, 80)

grade(s1)
grade(s2)
```


```{r}
class <- read.csv("student_homework.csv", row.names = 1)
```


```{r}
grade(class[1,])
```

APPLY grade() to the whole class

```{r}
ans <- apply(class, 1, grade)
```

```{r}
sort(ans, decreasing = TRUE)
```


```{r}
df1
df2
```

## One last function example


Find the intersection of two sets (two vectors)
# Start with a simple version of the problem

```{r}
x <- df1$IDs
y <- df2$IDs

intersect(x,y)
```


```{r}
x
y
x %in% y
which(x %in% y)
x[x %in% y]
```



```{r}
y %in% x
```


```{r}
gene_intersect <- function(x, y) {
  cbind(x[x %in% y], y[y %in% x])
}
```



# Improve by simplifying for human consumption

```{r}
gene_intersect4 <- function(df1, df2, gene.colname="IDs") {

 df1.name <- df1[,gene.colname]
 df2.name <- df2[,gene.colname]

 df1.inds <- df1.name %in% df2.name
 df2.inds <- df2.name %in% df1.name

 cbind( df1[ df1.inds, ],
 exp2=df2[ df2.inds, "exp"] )
}
```



# Improve by simplifying for human consumption
gene_intersect4 <- function(df1, df2, gene.colname="IDs") {

 df1.name <- df1[,gene.colname]
 df2.name <- df2[,gene.colname]

 df1.inds <- df1.name %in% df2.name
 df2.inds <- df2.name %in% df1.name

 cbind( df1[ df1.inds, ],
 exp2=df2[ df2.inds, "exp"] )
}
# Getting closer!
gene_intersect4(df1, df2)
#> IDs exp exp2
#> 2 gene2 1 -2
#> 3 gene3 1 1






# Test, break, fix, text again
df1 <- data.frame(IDs=c("gene1", "gene2", "gene3"),
 exp=c(2,1,1),
 stringsAsFactors=FALSE)
df3 <- data.frame(IDs=c("gene2", "gene2", "gene5", "gene5"),
 exp=c(-2, NA, 1, 2),
 stringsAsFactors=FALSE)
# Works but could do with more spit and polish!
gene_intersect4(df1, df3)
#> IDs exp exp2
#> 1 gene2 1 -2
#> 2 gene2 1 NA
#> Warning in data.frame(..., check.names = FALSE): row
names were found from a short variable and have been
discarded

```{r}
merge(df1, df2, by="IDs")
#> IDs exp.x exp.y
#> 1 gene2 1 -2
#> 2 gene3 1 1
```






